Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data

### 2018-10-02 paper.tex
Comments on chap 2 and 3

    \chapter{Performance}\label{performance}
        Neuroimaging frameworks are notoriously known to be compute-intensive.
        The tools and their respective workflows needed to incorporate 
        parallelism in order to save on computing time. Maintaining good 
        performance has become especially critical now that BigData has been
        introduced into the mix. The dichotomy remains between scientific 
        workflows and BigData workflows. While scientific workflows have made
        advances towards more performant engines, they still lack the 
        strategies BigData frameworks employ to reduce data transfers.

        Another notable aspect to consider is that both types of engines 
        (scientific and BigData) are tightly coupled to the infrastructure
        they were intended to be used on. That is, as scientific workflow 
        try to remain efficient in their handling of compute resources, as such
        resources were previously scarce. In contrast, BigData frameworks 
        make efficient use of commodity clusters, where data transfers are 
        burdensome, particularly when faced with BigData.

@@

        In this Chapter, we will compare scientific
        workflows to BigData frameworks on the basis of performance. 
        Section~\ref{fs} will look at what filesystems are required by the 
        engines. This is important because data movement of any kind is costly,
        and may be worse depending on the filesystems available. 
        Section~\ref{sched} will provide an overview of the schedulers 
        available and their compatibility with the workflow engines. 
        Section~\ref{other} will discuss other performance improvement 
        strategies employed by the workflow engines.

@@

            Unlike MapReduce and Spark-based workflows, Dask and scientific 
            workflows are not data-centric. They focus on the efficient 
            scheduling of tasks and rely on I/O and network transfers to
            communicate the data between nodes. With BigData workflows, such
            operations could incur significant performance penalties. 
            Filesystems. To mitigate this and remain efficient in the face of
            BigData, Dask implements lazy evaluation and in-memory processing.
            BigData platforms such as MapReduce rely heavily on I/O, but the 
            I/O performed occurs on the compute node rather than on 
            shared storaged, making it generally more efficient than scientific
            workflows. Additionally, MapReduce uses the concept of 
            locality to ensure limit data transfers across the network. Spark
            uses data locality, lazy evaluation and in-memory computing to 
            ensure its optimal performance.
            For the efficient processing of neuroimaging BigData, it will be
            necessary for scientific worflow engines to incorporate these 
            strategies.

@@

                    While Spark and Dask both rely on the concept of shared 
                    memory, Spark avoids data transfers as often as possible by
                    scheduling tasks to nodes where the memory is located. Dask
                    , however, will schedule its tasks to the next available 
                    node.

@@

                    In contrast to MapReduce, Spark does not require data to 
                    be stored on a HDFS-like filesystem. It can read data from
                    any form of shared-filesystem, including Lustre. If data 
                    is extracted from a non HDFS-like filesystem, the initial 
                    loading of data will not use data locality. However, in all
                    future tasks in which the data is already loaded into an 
                    RDD, the scheduler will be able to assign task to the most
                    favorable partition dictated by delay scheduling.

                \subsubsection{lazy evaluation}

                    Lazy evaluation is a strategy used throughout computer 
                    science to improve performance of operations. Rather than
                    evaluating a expression immediately, it will wait until it 
                    is required in order to compute the result. This is 
                    achieved in Spark through the calling of an \textit{action}
                    on an RDD and in Dask through the calling of a 
                    \textit{compute} function on one of its data structures.
                    By using lazy evaluation, the engines are capable of 
                    constructing graphs that result in the most optimal
                    task order of execution.

@@

            dask distributed
            \note{scientific: SLURM, PBS, SGE, HTCondor, Mesos (sometimes)
            big data: mapreduce - local or YARN, Spark - standalone, YARN or 
            Mesos, dask: any (easy to extend to any scheduler)}
        \section{Performance enhancement strategies}\label{other}
            scientific: workflow checkpointing (looking at file hashes)
            big data: data locality, in memory computing, lazy evaluation
    \chapter{Portability}\label{portability}
        In order for a workflow engine to be utilized, it much be simple for 
        users to install and execute the data. Platform-independence of both 
        the frameworks and tools, as well as, simple installation of all 
        necessary components and straightforward execution of workflows  are a 
        requirement for frameworks and their respective pipelines to be 
        portable. Without portability, users are limited to executing certain
        frameworks only or dealing with the significant overhead of making 
        these engines work on the system available to them. In this chapter, 
        we evaluate the portability of workflows on the basis of the following
        three components: 1) platform-independence, 2) tool installation and
        3) execution setup. We also examine containerization, one of the 
        strategies used to improve portability.

        \section{Platform-independence}
            Achieving platform-independence allows users to use the workflow 
            engine in any available environment. This however does not imply 
            that the tools, and as a consequence, the pipelines created will
            also be platform-independent. Many workflow engines have opted to 
            be platform independent. These workflow engines include: Nipype,
            PSOM, SPM, FASTR, LONI, AA, Pydpiper, Taverna, MapReduce, Spark and
            Dask. These workflow engines achieve platform independence through
            running on the Java Virtual Machine (JVM) or being written in 
            Python. Platform-independent engines that are Java-based include: 
            PSOM, SPM, LONI, AA, Taverna, MapReduce and Spark. The remainder of
            the pipelines are implemented in Python. The remainder of the 
            workflow engines (Pegasus, Make, OpenMOLE, Nextflow, Galaxy, and 
            Toil are not platform-independent, but can execute on UNIX-like 
            systems.

            Engines built ontop of a platform-independent language does not 
            necessarily imply platform independence. For instance, OpenMOLE is 
            built in Scala and Netflow runs on Groovy, yet neither are platform
            independent. In the case of OpenMOLE, this is due to its CARETask
            blackbox, which is only possible for Linux environments. However,
            many tools and programs available for incorporation into workflows
            are platform-dependent. For example, even though Nipype is 
            platform-independent, common tools used in Nipype modules, such as 
            FreeSurfer and FSL, are Linux-based. Moreover, HPC infrastructure 
            is usually Linux-based and it is therefore crucial that all 
            workflows are compatible with and tested on Linux systems.

@@

            Unless a user is building a workflow exclusively based on custom
            functions or using a monolithic workflow engine such as SPM, they
            will need to install the tools necessary for their workflow on
            their computer or cluster. Depending on the workflow engine 
            selected, this may be managed by the engine, or the user may be 
            left to install the dependencies themselves. Workflow engines that
            delegate tool installation to the user include: Nipype, PSOM,

>>>>

    \chapter{Performance}\label{performance}
        Neuroimaging frameworks are notoriously known to be compute-intensive.
        The tools and their respective workflows needed to incorporate 
        parallelism in order to save on computing time. Maintaining good 
        performance has become especially critical now that BigData has been
        introduced \weird{into the mix}. The dichotomy remains between scientific 
        workflows and BigData workflows. While scientific workflows have made
        advances towards more performant engines, they still lack the 
        strategies BigData frameworks employ to reduce data transfers.

        Another notable aspect to consider is that both types of engines 
        (scientific and BigData) are tightly coupled to the infrastructure
        they were intended to be used on. That is, as scientific workflows \tristan{use 'engine' and 'workflows' consistently (think nipype vs fmriprep).}
        try to remain efficient in their handling of compute resources, as such
        resources were previously scarce. In contrast, BigData frameworks 
        make efficient use of commodity clusters, where data transfers are 
        burdensome, particularly when faced with BigData.

@@

        In this Chapter, we will compare scientific
        workflows to BigData frameworks on the basis of performance. 
        Section~\ref{fs} will look at what filesystems are \weird{required} \tristan{supported?} by the 
        engines. This is important because data movement of any kind is costly,
        and may be worse depending on the filesystems available. \tristan{Also, mention that neuroinformatics
        pipelines manipulate files.}
        Section~\ref{sched} will provide an overview of the schedulers 
        available and their compatibility with the workflow engines. 
        Section~\ref{other} will discuss other performance improvement 
        strategies employed by the workflow engines.

@@

            Unlike MapReduce and Spark-based workflows, Dask and scientific 
            workflows are not data-centric. They focus on the efficient 
            scheduling of tasks and rely on I/O and network transfers to
            communicate the data between nodes. With BigData workflows, such
            operations could incur significant performance penalties. 
            \weird{Filesystems}. To mitigate this and remain efficient in the face of
            BigData, Dask implements lazy evaluation and in-memory processing.
            BigData platforms such as MapReduce rely heavily on I/O, but the 
            I/O performed occurs on the compute node rather than on 
            shared storaged, making it generally more efficient than scientific
            workflows. Additionally, MapReduce uses the concept of 
            locality to ensure limit data transfers across the network. Spark
            uses data locality, lazy evaluation and in-memory computing to 
            ensure its optimal performance.
            For the efficient processing of neuroimaging BigData, it will be
            necessary for scientific worflow engines to incorporate these 
            strategies.

@@

                    While Spark and Dask both rely on the concept of shared 
                    memory, Spark avoids data transfers as often as possible by
                    scheduling tasks to nodes where the memory is located \tristan{You mean where RAM is available or where the data is located?}. Dask
                    , however, will schedule its tasks to the next available 
                    node.

@@

                    In contrast to MapReduce, Spark does not require data to 
                    be stored on a HDFS-like filesystem. It can read data from
                    any form of shared-filesystem, including Lustre. If data 
                    is extracted from a non HDFS-like filesystem, the initial 
                    loading of data will not use data locality. However, in all
                    future tasks in which the data is already loaded into an 
                    RDD, the scheduler will be able to assign task to the most
                    favorable partition dictated by delay scheduling \tristan{ignoring data locality?}.

                \subsubsection{Lazy evaluation}

                    Lazy evaluation is a strategy used throughout computer 
                    science to improve performance of operations. Rather than
                    evaluating a expression immediately, it will wait until it 
                    is required in order to compute the result. This is 
                    achieved in Spark through the calling of an \textit{action}
                    on an RDD and in Dask through the calling of a 
                    \textit{compute} function on one of its data structures.
                    By using lazy evaluation, the engines are capable of 
                    constructing graphs that result in the most optimal
                    task order of execution. \tristan{And further save on data transfers}

@@

            dask distributed
            \note{scientific: SLURM, PBS, SGE, HTCondor, Mesos (sometimes)
            big data: mapreduce - local or YARN, Spark - standalone, YARN or 
            Mesos, dask: any (easy to extend to any scheduler)}
        \section{Performance enhancement strategies}\label{other}
            scientific: workflow checkpointing (looking at file hashes)
            big data: data locality, in memory computing, lazy evaluation

            \tristan{Mentioning fault tolerance would be wise too.}

    \chapter{Portability}\label{portability}
        In order for a workflow engine to be utilized, \weird{it much be simple} for 
        users to \weird{install and execute the data}. Platform-independence of both 
        the frameworks and tools, as well as, simple installation of all 
        necessary components and straightforward execution of workflows  are a 
        requirement for frameworks and their respective pipelines to be 
        portable. Without portability, users are limited to executing certain
        frameworks only or dealing with the significant overhead of making 
        these engines work on the system available to them. In this chapter, 
        we evaluate the portability of workflows on the basis of the following
        three components: 1) platform-independence, 2) tool installation and
        3) execution setup. We also examine containerization, one of the 
        strategies used to improve portability.

        \section{Platform-independence}
            Achieving platform-independence allows users to use the workflow 
            engine in any available environment. This however does not imply 
            that the tools, and as a consequence, the pipelines created will
            also be platform-independent. Many workflow engines have opted to 
            be platform independent. These workflow engines include: Nipype,
            PSOM, SPM, FASTR, LONI, AA, Pydpiper, Taverna, MapReduce, Spark and
            Dask. These workflow engines achieve platform independence through
            running on the Java Virtual Machine (JVM) or being written in 
            Python. Platform-independent engines that are Java-based include: 
            PSOM, SPM, LONI, AA, Taverna, MapReduce and Spark \tristan{this list contains Matlab
            engines}. The remainder of
            the pipelines are implemented in Python. The remainder of the 
            workflow engines (Pegasus, Make, OpenMOLE, Nextflow, Galaxy, and 
            Toil are not platform-independent, but can execute on UNIX-like 
            systems \tristan{macOS and Linux}.

            Engines built on top of a platform-independent language does not 
            necessarily imply platform independence \tristan{Grammar? Engines ... does not}. For instance, OpenMOLE is 
            built in Scala and Netflow runs on Groovy, yet neither are platform
            independent. In the case of OpenMOLE, this is due to its CARETask
            blackbox, which is only possible for Linux environments. However,
            many tools and programs available for incorporation into workflows
            are platform-dependent. For example, even though Nipype is 
            platform-independent, common tools used in Nipype modules, such as 
            FreeSurfer and FSL, are Linux-based. Moreover, HPC infrastructure 
            is usually Linux-based and it is therefore crucial that all 
            workflows are compatible with and tested on Linux systems.

@@

            Unless a user is building a workflow exclusively based on custom
            functions or using a monolithic workflow engine such as SPM, they
            will need to install the tools necessary for their workflow on
            their computer or cluster \tristan{replace computer or cluster by infrastructure?}. Depending on the workflow engine 
            selected, this may be managed by the engine, or the user may be 
            left to install the dependencies themselves. Workflow engines that
            delegate tool installation to the user include: Nipype, PSOM,
